Forecasting the 2024 U.S. Presidential Election

### 2024-11-04 paper/paper.qmd
Update paper.qmd
Added survey questions and refined appendix.

Our analysis and prediction utilizes data collected by a single pollster and it is important to scrutinize the sampling methodology behind the sampling methods used to determine the extent of the external validity of our predictions. Primarily, the data gathered by New York Times was done so with the collaboration with Siena College through live-interviewers reaching voters by telephone. 
The poll targets registered voters in the United States, with a particular focus on voters in swing states for presidential races. Consequently, the target population is registered voters of the United States of America and the sampling frame is the national list of registered voters from which NYC employs random sampling to create their pollsters. And from this sampling frame, respondents are randomly selected while balancing key demographic and political characteristics through both front-end sampling and post-survey weighting pursuant to appropriate representation of the entire target population. 
The poll relies on random digit dialing to reach voters, predominantly via cell phones. While landlines are also called, the dominance of cellphone usage aligns with contemporary communication trends. This results in their rendition of random sampling, which contributes to the external validity of data. However, though the sample was selected randomly, reaching voters by phone has become more difficult due to low response rates requiring many attempts to secure responses from certain individuals (response rates are as  low as 2% at NYC). Despite this, the phone methodology is still seen as effective for its ability to quickly reach a diverse voter pool across various states as well as seen as the golden standard for polling. Although this can be seen as random sampling, the concern of low response rates and reliance on dialing to reach voters raises concerns of selection bias. There are questions as to whether a certain demographic will be more responsive to being reached through a cell phone. Moreover, voters that are reached over the phone may feel the urge to provide more ‘socially desirable’ responses that are not reflective of their true opinions. 
With a response rate of 2%, there is valid concern over non-response bias underrepresenting certain demographics. The poll addresses non-response through a combination of front-end sampling adjustments (calling harder-to-reach voters) and post-survey weighting. Weighting is applied to ensure that the final sample aligns with the demographic and political composition of the electorate, based on voter file data. Non-response bias is a persistent challenge, and even though weighting helps, it's impossible to be certain that respondents are fully representative of those who do not respond. Experiments with alternative methods, such as mail surveys with incentives of cash vouchers have shown to increase response rate up to 30% (New York Times).

@@

##  Methodology
The Times/Siena Poll employs a stratified random sampling approach to ensure that the sample accurately represents the diversity within the target population of registered voters in the United States. This method involves dividing the population into distinct subgroups, or “strata,” based on key characteristics such as state, age, gender, race, education level, and political affiliation. For example, in presidential elections, particular emphasis is placed on voters from swing states like Arizona, North Carolina, Georgia, and Pennsylvania. By focusing on these battleground states, the poll aims to capture the opinions of voters who are most likely to influence the election outcome. Once the strata are established, respondents are randomly selected within each group, ensuring that all demographic groups are appropriately represented in proportion to their actual distribution within the electorate.

## Advantages
Stratified sampling offers the advantage of higher precision in representing the population's diversity, as each subgroup is proportionally represented. By breaking down the sample into distinct groups, the Times/Siena Poll can adjust for demographic variations between states or voter groups, which is particularly useful in a diverse electorate. This approach helps to minimize sampling error by ensuring that all significant voter subgroups are included, thus increasing the reliability of the poll’s predictions.

@@

Although stratified sampling improves representativeness, it presents certain challenges in execution. First, identifying and reaching a sufficient number of respondents from harder-to-reach groups (such as young voters, minority populations, and those without reliable phone access) can be difficult, especially given the 2% response rate. To counterbalance the low response rates, the poll requires multiple call attempts and careful post-survey weighting to adjust for any potential underrepresentation. Additionally, while phone polling offers broad geographic reach and a standardized mode of data collection, it limits the sample to voters with phone access who are willing to participate in phone interviews, potentially introducing selection bias if certain demographics are less reachable by phone.

@@

According to the New York Times, with a sample of 1000 individual votes, there is a margin of error of around three to four percentages. With $100,000, there are limitations on the labor and resources that is required to poll 1000 voters while balancing out demographics through dialing. Furthermore, there is additional filtering of our sample frame to determine likely voters. Assuming New York Times’ 2% voter response rate, to sample 1000 voters, at least 50,000 voters will need to be called. It is likely that our team will be dialing way more than 50,000 times since it is an unreasonable assumption that every voter we call would pick up or be available to provide a response. There is also the fact that our team does not carry the same credibility and history of acquiring poll data as the New York Times does, resulting in more reluctance for voters to share information with us.
Therefore, it is not feasible for our methodology to employ telephone dialing as our method to reach voters. Instead, we focus on online panels and surveys with incentives for completing the survey correctly. However, online panels and surveys come with their own set of challenges and weaknesses that must be addressed in our methodology. Primarily, the problem of self-selection in online panels disrupts the random selection that allows the external validity of our predictions. Moreover, there is apossibility that we acquire responses from respondents that are not likely to vote, gaining data from outside our target population. Even if we distribute the panels randomly among swing states, responses are still corrupted by selection bias that occurs from responses that are provided from demographics that feel more passionately about the election or are more likely to respond to online panels. Unlike telephone dialing, we have less control over the specific demographic we want to target, resulting in a less stratified data that is representative of the target population overall.

@@

Although weighting adjustments helps  us represent our  target population more closely, there is nothing to stop respondents from outside our target population to participate in our panels. To produce accurate predictions of election results, our sample needs to be acquired from a target population that will participate in the election. Therefore, it becomes necessary to include screening questions designed to filter for likely voters, based on voting history or engagement levels. Our survey includes questions that pertain to voting history, interest and engagement, voting intention, and eligibility. Responses to these questions will be used to produce an aggregate “likely voter” score that filters for the target population enabling for more accurate predictions of the presidential election.

##  Elements of Our Survey
To gather accurate responses from a diverse and representative set of likely voters, our survey design carefully adheres to best practices in methodology. Given our goal to avoid attrition bias, we’ve restricted the estimated survey completion time to under ten minutes. This constraint is stricter than the fifteen-minute benchmark typically used by NYC and Siena College for telephone surveys, as online surveys require shorter durations to maintain respondent engagement. Our survey is divided into three main sections: Preamble, Eligibility, and Party Affiliation. The survey we constructed can be found in Appendix B.

@@

@@

>>>>

Our analysis and prediction utilizes data collected by a single pollster and it is important to scrutinize the sampling methodology behind the sampling methods used to determine the extent of the external validity of our predictions. Primarily, the data gathered by New York Times was done so with the collaboration with Sienna College through live-interviewers reaching voters by telephone.

## Target Population and Sampling Frame
The poll targets registered voters in the United States, with a particular focus on voters in swing states for presidential races. Consequently, the target population is registered voters of the United States of America and the sampling frame is the voter file, a national list of registered voters from which NYC employs random sampling to create their pollsters. This voter file contains information on more than 200 million Americans that includes information about their party affiliation, voting history, and  basic demographic information. From this sampling frame, respondents are randomly selected while balancing key demographic and political characteristics through both front-end sampling and post-survey weighting pursuant to appropriate representation of the entire target population.

## Recruitment Methods
The poll relies on random digit dialing to reach voters, predominantly via cell phones. While landlines are also called, the dominance of cellphone usage aligns with contemporary communication trends. This results in their rendition of random sampling, which contributes to the external validity of data. However, reaching voters by phone has become more difficult due to low response rates requiring many attempts to secure responses from certain individuals (response rates are as  low as 2% at NYC). In fact, according to an article by Emerson College, telephone polling has become entirely obsolete (@EmersonCollege). Despite this, the phone methodology is still seen as effective for its ability to quickly reach a diverse voter pool across various states as well as seen as the golden standard for polling. Although this can be seen as random sampling, the concern of low response rates and reliance on dialing to reach voters raises concerns of selection bias. There are questions as to whether a certain demographic will be more responsive to being reached through a cell phone. Moreover, voters that are reached over the phone may feel the urge to provide more ‘socially desirable’ responses that are not reflective of their true opinions.

## Non-Response Remedies
With a response rate of 2%, there is valid concern over non-response bias underrepresenting certain demographics. The poll addresses non-response through a combination of front-end sampling adjustments (calling harder-to-reach voters) and post-survey weighting. Weighting is applied to ensure that the final sample aligns with the demographic and political composition of the electorate, based on voter file data. Non-response bias is a persistent challenge, and even though weighting helps, it's impossible to be certain that respondents are fully representative of those who do not respond. Experiments with alternative methods, such as mail surveys with incentives of cash vouchers have shown to increase response rate up to 30% (New York Times).

@@

##  Methodology
The Times/Siena Poll employs a stratified random sampling approach to ensure that the sample accurately represents the diversity within the target population of registered voters in the United States. This method involves dividing the population into distinct subgroups, or “strata,” based on key characteristics such as state, age, gender, race, education level, and political affiliation. For example, in presidential elections, particular emphasis is placed on voters from swing states like Arizona, North Carolina, Georgia, and Pennsylvania. By focusing on these battleground states, the poll aims to capture the opinions of voters who are most likely to influence the election outcome. Once the strata are established, respondents are randomly selected within each group, ensuring that all demographic groups are appropriately represented in proportion to their actual distribution within the electorate. Furthermore, post-survey weighting is employed to more closely resemble this distribution.

## Advantages
Stratified sampling offers the advantage of higher precision in representing the population's diversity, as each subgroup is proportionally represented. By breaking down the sample into distinct groups, the Times/Siena Poll can adjust for demographic variations between states or voter groups, which is particularly useful in a diverse electorate. This approach helps to minimize sampling error by ensuring that all significant voter subgroups are included, thus increasing the reliability of the poll’s predictions. It is standard practice for them to sample 1,000 respondents for their polls. Using stratified sampling to have this sample size closely resemble that of the entire target population cuts costs and provides efficient real-time data.

## Stratified Sampling over Random Sampling

If the methodology randomly sampled from the target population without regard for the distribution of demographics in the state, significant increases to the sample sizes have to be made. Random sampling will have largely less precision in representing the distribution of different demographics within a given state unless a large enough sample size were to be drawn. Although it is ideal to have probability based sampling increasingly resemble the demographic distribution quota with growing sample sizes, it is unfeasible to conduct such a large scale poll as this approach would encounter a limitation of resources. With declining response rates and the difficulty of reaching voters, New York Times/Siena's nonprobability sampling is a more realistic and efficient method of polling than probability sampling is (@tellingstorieswithdata). As such, the stratified sampling approach allows New York Times/Siena to represent the entire target population with a sample size of 1,000.

@@

Although stratified sampling improves representativeness, it presents certain challenges in execution. First, identifying and reaching a sufficient number of respondents from harder-to-reach groups (such as young voters, minority populations, and those without reliable phone access) can be difficult, especially given the 2% response rate. To counterbalance the low response rates, the poll requires multiple call attempts and careful post-survey weighting to adjust for any potential underrepresentation. Additionally, while phone polling offers broad geographic reach and a standardized mode of data collection, it limits the sample to voters with phone access who are willing to participate in phone interviews, potentially introducing selection bias if certain demographics are less reachable by phone. Although New York Times/Sienna College employs random digit dialing from their sampling frame, their rendition of sample recruitment requires them to sample individuals according to quotas, resulting in nonprobability sampling.

## Evaluation of Questionnaire Design in the Times/Siena Poll

### Strengths

The Times/Siena Poll’s approach to questionnaire design reflects a careful effort to craft unbiased, relevant, and inclusive questions. By tailoring questions to current events and topics of widespread interest, the poll aims to capture public sentiment on issues that directly affect voters' attitudes, such as economic conditions or recent political debates. This responsiveness to current events strengthens the poll's relevance and allows it to track shifts in opinion over time. The poll’s commitment to neutrality is a significant strength. Each question is designed to ensure that respondents, regardless of their political views, feel that their perspectives are fairly represented. The team behind the poll spends substantial time crafting response options to reflect a balanced spectrum of views, minimizing potential response bias and enhancing the questionnaire's reliability across different demographic and ideological groups (@NewYorkTimes). Clarity is also a central focus in their questionnaire development. The team emphasizes ensuring that each question is clear and unambiguous, minimizing the risk of misinterpretation. This dedication to straightforward, widely understandable questions improves the validity of responses, as it reduces the chance that respondents interpret questions differently based on background or knowledge level.

### Limitations

Despite the strengths of the Times/Siena Poll’s questionnaire design, there are some limitations. One notable concern is that, although the poll tries to avoid leading questions, any standardized questionnaire may still unintentionally prime respondents or introduce subtle biases, especially when questions involve sensitive or polarizing topics. Although the pollster tries to avoid attrition bias by keeping each interview less than 15 minutes, in seeking to make questions clear and concise, the questionnaire may not capture the full nuance of voter opinions, especially on multifaceted issues like economic policy or social justice. While the poll strives to capture real views without introducing new ideas, respondents are still prone to a social desirability bias. This is especially possible in phone interviews, where respondents may be more inclined to provide answers they perceive as socially acceptable rather than expressing their true opinions. Although the design aims to be fair, the influence of an interviewer’s presence may subtly affect responses (@thepsygologyofsurveyresponse)

@@

According to the New York Times, with a sample of 1000 individual votes, there is a margin of error of around three to four percentages. With $100,000, there are limitations on the labor and resources that is required to poll 1000 voters while balancing out demographics through dialing. Furthermore, there is additional filtering of our sample frame to determine likely voters. Assuming New York Times’ 2% voter response rate, to sample 1000 voters, at least 50,000 voters will need to be called. It is likely that our team will be dialing abundantly more than 50,000 times since it is an unreasonable assumption that every voter we call would pick up or be available to provide a response. There is also the fact that our team does not carry the same credibility and history of acquiring poll data as the New York Times does, resulting in more reluctance for voters to share information with us.
Therefore, it is not feasible for our methodology to employ telephone dialing as our method to reach voters. Instead, we focus on online panels and surveys with incentives for completing the survey correctly. However, online panels and surveys come with their own set of challenges and weaknesses that must be addressed in our methodology. Primarily, the problem of self-selection in online panels disrupts the random selection that allows the external validity of our predictions. Moreover, there is apossibility that we acquire responses from respondents that are not likely to vote, gaining data from outside our target population. Even if we distribute the panels randomly among swing states, responses are still corrupted by selection bias that occurs from responses that are provided from demographics that feel more passionately about the election or are more likely to respond to online panels. Unlike telephone dialing, we have less control over the specific demographic we want to target, resulting in a less stratified data that is representative of the target population overall.

@@

Although weighting adjustments helpsus represent our target population more closely, there is nothing to stop respondents from outside our target population to participate in our panels since it will be advertised online. To produce accurate predictions of election results, our sample needs to be acquired from a target population that will participate in the election. Therefore, it becomes necessary to include screening questions designed to filter for likely voters, based on voting history or engagement levels. Our survey includes questions that pertain to voting history, interest and engagement, voting intention, and eligibility. Responses to these questions will be used to produce an aggregate “likely voter” score that filters for the target population enabling for more accurate predictions of the presidential election.

##  Elements of Our Survey
To gather accurate responses from a diverse and representative set of likely voters, our survey design carefully adheres to best practices in methodology. Given our goal to avoid attrition bias, we’ve restricted the estimated survey completion time to under ten minutes. This constraint is stricter than the fifteen-minute benchmark typically used by NYC and Siena College for telephone surveys, as online surveys require shorter durations to maintain respondent engagement. Our survey is divided into three main sections: Preamble, Eligibility, and Party Affiliation.

@@

# Budget Allocation
  Advertisements for sampling recruitment: $70,000
  Prize Pool for Incentive: $5,000
  Data Processing and Cleaning: $15,000
  Miscellaneous Expenses: $5,000
  Petty Cash: $5,000

@@

# Structure

## Preamble

Title: 2024 Presidential Election Poll Data

Introduction: Thank you for your participation in our $5,000 giveaway program.

For entry, please complete the survey in its entirety below. Please be informed that your answers are strictly confidential and will, under no circumstances, be shared with any party, group, or persons of affiliation.

$5,000 Cash Voucher Program Details: Completing the survey below qualifies as one entry into our giveaway. Each participant receives a unique entry identifier, and every entry will have an equal chance in our final random selection of the prize winner. Please note that each individual may complete only one survey.

Referral Program: Our referral program offers you the chance to increase your odds of winning. After completing the survey, you’ll receive a unique referral code. If a friend or associate completes the survey using your code, your total number of entries doubles, effectively doubling your chances of winning the $5,000 prize.

Finally, we wish to clarify that this survey is conducted purely for research purposes and is entirely unaffiliated with any political party, organization, or individual in a position of power. Our goal is to gain insight into the evolving political landscape and levels of public engagement.

Referral Code (please ignore if inapplicable)

If you have entered a referral code above, please indicate the relationship with the referee from the best choice below.
Friend
Associate
Relative
Other…

## Eligibility
Do you meet your state's residency requirements?
Yes
No (End Survey)
Please select "N/A" from the options below if you reside in North Dakota. 
Which of the following best describes you?
I am registered to vote
I am not registered to vote
I plan to register
N/A
Full Name (For identification of $5,000 cash winner)
From which state will you be vote for the next president of the United States of America?
Where can we reach you when we announce the winner?
E-mail
Telephone
Other…
Contact information
**We will not use this for any other purpose than to announce winners. It will only be used to reach our lucky winner for instructions on receiving the cash prize.
Which of the following age groups best describes you?
18-25
26-31
32-40
41-50
51-60
61-80
81-125
Under 18
Which of the following best describes you?
American Indian or Alaskan Native
Asian
African American
Hispanic or Latino
Native Hawaiian or Pacific Islander
White
Middle Eastern or North African
Other…
What is your gender?
Male
Female
Prefer not to say
Other…
Which of the following situations best describe you?
I did not vote in the 2020 presidential election
I planned to vote but wasn’t able to
I voted in the 2020 presidential election
How likely are you to vote in the upcoming election?
Very unlikely
1
2
3
Very likely
Have you attended a presidential rally or their campaigns this year?
Yes
No
Only in the past
On a scale of 1 to 3, how closely do you follow the presidential race?
Not at all
1
2
3
4
5
Very closely

## Party Affiliation
Which party do you prefer a candidate to be chosen from?
Democratic Party's candidate
Republican Party's candidate
Another Party's candidate
In the 2020 presidential election, who did you vote for?  If you did not vote, who would you have voted for?
Donald Trump
Joe Biden
Neither
Other…
If planning to vote, which presidential candidate are you planning to vote for?
If you are abstaining, please select which candidate you feel more strongly about.
Donald Trump
Kamala Harris
Other...
\newpage
